---
title: "100_Most_Observed_sp"
author: "Thomas Schneider--Bruchon"
date: "01/04/2020"
output: html_document
---

# Recovery of the 100 most observed species and passage to the square cordinates (!= points)

# Environnement 
```{r, message=FALSE}
rm(list=ls())
library(sf)
library(maps)
library(dplyr)
library(units)
library(ggplot2)
library(lubridate)
```


# Variables
```{r}
path = "/Users/Schneider/STAGE_STOC_MNHN/Donnees"
load(paste(path,"/observation_STOC.Rda", sep = ""))
load(paste(path,"/inventaire_STOC.Rda", sep = ""))
```

# Recovery of 100 most abundant species
```{r}
selection <- select(observation_STOC, code_sp,abondance)
datatable <- aggregate(selection$abondance, by = list(code_sp=selection$code_sp), FUN = sum)
order_sp <- datatable[order(datatable$x, decreasing = T),]
ab_sp <- order_sp[1:101,]
```


# Get all the data for these species
```{r}
try(sp_data <- observation_STOC %>%
filter(espece %in% ab_sp$code_sp), silent = T)

sp_data <- select(sp_data,id_inventaire,id_carre,date,code_sp,abondance,distance_contact)
date <- year(as.Date(sp_data$date))
sp_data$date <- date
```


# Cleaning
```{r}
rm(observation_STOC,ab_sp,datatable,order_sp,selection,date)
```


# Recovery year passing
```{r}
id_inventaire <- sp_data$id_inventaire

try(selection <- inventaire_STOC %>%
filter(pk_inventaire %in% id_inventaire), silent = T)

id_inventaire <- as.character(selection$pk_inventaire)
passage_ob <- selection$passage_stoc

inv_data <- as.data.frame(cbind(id_inventaire,passage_ob))
```


#Cleaning
```{r}
rm(inventaire_STOC,selection,id_inventaire,passage_ob)
```


# Delete contact distance "transit", date < 2001, and passage "NA"
```{r}
datatable <- inner_join(inv_data,sp_data)
datatable <- subset(datatable,distance_contact != "TRANSIT")
datatable <- subset(datatable,date > "2001")
datatable <- subset(datatable,passage_ob != "NA")
```

# Sum of species abundance per square, year, passage
```{r}
datatable <- aggregate(datatable$abondance, by = list(id_carre=datatable$id_carre, code_sp=datatable$code_sp, date=datatable$date, passage=datatable$passage_ob), FUN = sum)
colnames(datatable)[5] <- "abondance"
```


# Recovery of square coordinates
```{r}
dataframe <- read.csv(paste(path,"/carre_2019-03-13.csv", sep=""), header=T, sep=";",  encoding="UTF-8")
selection <- select(dataframe, pk_carre, longitude_grid_wgs84, latitude_grid_wgs84)

obs_STOC_100sp <- inner_join(datatable,selection, by = c("id_carre" = "pk_carre"))
```


#Cleaning
```{r}
rm(dataframe,selection,datatable,sp_data,inv_data)
```


# Save
```{r}
save_path = "/Users/Schneider/STAGE_STOC_MNHN/Donnees/"
save(obs_STOC_100sp, file = paste(save_path,"/obs_STOC_100sp.Rda", sep = ""))
```


