---
title: "All_species_dataset"
author: "Thomas Schneider--Bruchon"
date: "16/04/2020"
output: html_document
---

# Environment
```{r}
rm(list=ls())
library(sf)
library(dplyr)
library(units)
library(lubridate)
```


# Variables
```{r}
path = "/Users/Schneider/STAGE_STOC_MNHN/Donnees"
load(paste(path,"/observation_STOC.Rda", sep = ""))
load(paste(path,"/espece_BL_STOC.Rda", sep = ""))
load(paste(path,"/BL_all_polygons_F_remanie.Rda", sep = ""))
```

# Recovery year passing
```{r}
date <- year(as.Date(observation_STOC$date))
observation_STOC$date <- date

load(paste(path,"/inventaire_STOC.Rda", sep = ""))
id_inventaire <- observation_STOC$id_inventaire

try(selection <- inventaire_STOC %>%
filter(pk_inventaire %in% id_inventaire), silent = T)
id_inventaire <- as.character(selection$pk_inventaire)
passage_ob <- selection$passage_stoc
inv_data <- as.data.frame(cbind(id_inventaire,passage_ob))

#Cleaning 
rm(inventaire_STOC,selection,id_inventaire,passage_ob,date)
```


# Delete contact distance "transit", date < 2001, and passage "NA"
```{r}
datatable <- inner_join(inv_data,observation_STOC)
datatable <- subset(datatable,distance_contact != "TRANSIT")
datatable <- subset(datatable,date > "2001")
datatable <- subset(datatable,passage_ob != "NA")
```


# Sum of species abundance per square, year, passage
```{r}
datatable <- aggregate(datatable$abondance, by = list(id_carre=datatable$id_carre, code_sp=datatable$code_sp, date=datatable$date, passage=datatable$passage_ob), FUN = sum)
colnames(datatable)[5] <- "abondance"
```


# Recovery of square coordinates
```{r}
dataframe <- read.csv(paste(path,"/carre_2019-03-13.csv", sep=""), header=T, sep=";",  encoding="UTF-8")
selection <- select(dataframe, pk_carre, longitude_grid_wgs84, latitude_grid_wgs84)

all_obs_STOC <- inner_join(datatable,selection, by = c("id_carre" = "pk_carre"))
```


# Cleaning 
```{r}
rm(dataframe,selection,datatable,inv_data,observation_STOC)
```


# Save
```{r}
save_path = "/Users/Schneider/STAGE_STOC_MNHN/Donnees/"
save(all_obs_STOC, file = paste(save_path,"/all_obs_STOC.Rda", sep = ""))
```


# If you have already do this part you can just load the "all_obs_STOC" datatable
```{r}
load(paste(path,"/all_obs_STOC.Rda", sep = ""))
```


# Get the BL area for 1 species
```{r}
area <- function(code_sp, comparaison_table, area_table){
  
    #get the species BL name
    try(row <- comparaison_table %>%
    filter(pk_species %in% code_sp), silent = T)
    species <- as.character(row$SCINAME)
    cat(species,"\n")
    
    #get the species distribution area
    try(area <- area_table %>%
    filter(SCINAME %in% species), silent = T)
    
    if(length(area)<1){
      return(paste("Error,",species,"n'est pas dans la table de comparaison"))
    }
  
    #get only resident and breeding season
    try(good_area <- area %>%
    filter(SEASONAL %in% c(1,2)), silent = T)
  
    return(good_area)
}
```



# Get the observations for 1 species
```{r}
observation <- function(sp_code, observation_table){
    
    #get the observations
    try(obs_sp <- observation_table %>%
    filter(code_sp %in% sp_code), silent = T)
    
    if(length(sp_code)<1){
      return(paste("Error,",sp_code,"n'est pas dans la table d'observation"))
    }
    
    obs_sp <- subset(obs_sp, longitude_grid_wgs84 != "NA")
    
    return(obs_sp)
}
```




# Get id_carre
```{r}
get_carre <- function(observation_species){
    
    carre <- select(observation_species, id_carre, longitude_grid_wgs84, latitude_grid_wgs84)
    carre <- distinct(carre)

    return(carre)
}
```




# Function:
# Bool == 1  ->   return(POINT IN POLYGON)
# Bool == 0  ->   return(POINT EXT POLYGON)
```{r}
position_car_pol <- function(obs_sp,area_sp,bool){
  
  # Conversion to Lambert coordinates
  polygon_Lambert <- st_transform(area_sp,2154)
  
  #Initialisation
  point_dans_polygon <- c()
  point_ext_polygon <- c()
  
  #print("START PHASE 1/2")
  
  for(i in 1:nrow(obs_sp)){
    
    cordinates <- st_point(c(obs_sp[i,]$longitude_grid_wgs84, obs_sp[i,]$latitude_grid_wgs84))
    cordinates <- st_sfc(cordinates)
    st_crs(cordinates) <- 4326
    cordinates <- st_transform(cordinates,2154)
    
    if(bool > 0){
      if(st_within(cordinates,polygon_Lambert) %>% lengths > 0){
        row = st_sf(id_carre=obs_sp[i,]$id_carre, cor_carre=cordinates)
        point_dans_polygon <- rbind(point_dans_polygon,row)
     }
    }
    else{
      if(!st_within(cordinates,polygon_Lambert) %>% lengths > 0){
        row = st_sf(id_carre=obs_sp[i,]$id_carre, cor_carre=cordinates)
        point_ext_polygon <- rbind(point_ext_polygon,row)
      }
    }
    
    if(i%%(ceiling(nrow(obs_sp)/20)) == 0  || i == nrow(obs_sp)){
      #print(paste(round((i/nrow(obs_sp))*100, 0),"%", sep =""))
      if(bool==0){
        save(point_ext_polygon, file = paste("Test_",ceiling(i/ceiling(nrow(obs_sp)/20)),".Rda",sep=""))
        rm(point_ext_polygon)
        point_ext_polygon <- c()
        j = ceiling(i/ceiling(nrow(obs_sp)/20)) 
      }
      else{
        save(point_dans_polygon, file = paste("Test_",ceiling(i/ceiling(nrow(obs_sp)/20)),".Rda",sep=""))
        rm(point_dans_polygon)
        point_dans_polygon <- c()
        j = ceiling(i/ceiling(nrow(obs_sp)/20)) 
      }
    }
  }
  
  
  #print("END PHASE 1/2")
  
  rlist <- c()
  #print("START PHASE 2/2")
  
  point_ext_polygon_f <- c()
  point_dans_polygon_f <- c()
  
  
  for(i in 1:j){
    #print(paste(round((i/j)*100, 0),"%", sep =""))
    load(paste("Test_",i,".Rda",sep=""))
    rlist <- c(rlist,(paste("Test_",i,".Rda",sep="")))
    if(bool == 0){
      dataset = point_ext_polygon
      point_ext_polygon_f = rbind(point_ext_polygon_f,dataset)
      rm(dataset)
    }
    else{
      dataset = point_dans_polygon
      point_dans_polygon_f = rbind(point_dans_polygon_f,dataset)
      rm(dataset)
    }
  }

  
  #print("END PHASE 2/2")
  file.remove(rlist)
  rm(rlist)
  
  # Return of dataframes
  if(bool == 0){
    if(!is.null(nrow(point_ext_polygon_f))){
      EXT <- rep(0, nrow(point_ext_polygon_f))
      point_ext_polygon_f <- cbind(point_ext_polygon_f,EXT)
      colnames(point_ext_polygon_f)[2] <- "INT/EXT"
      return(point_ext_polygon_f)
    }
    else{
      dataframe_vide <- c()
      return(dataframe_vide)
    }
    
  }
  else{
    if(!is.null(nrow(point_dans_polygon_f))){
      INT <- rep(1, nrow(point_dans_polygon_f))
      point_dans_polygon_f <- cbind(point_dans_polygon_f,INT)
      colnames(point_dans_polygon_f)[2] <- "INT/EXT"
      return(point_dans_polygon_f)
    }
    else{
      dataframe_vide <- c()
      return(dataframe_vide)
    }
  }
}
```



# Function
# Get the closest point of the nearest polygon for an EXT observation point
```{r}
nearest_points_ext <- function(cor_carre_obs,area_sp){
  
  if(is.null(nrow(cor_carre_obs))){
    dataframe_vide <- c()
    return(dataframe_vide)
  }
  
  # Conversion to Lambert coordinates
  polygon_Lambert <- st_transform(area_sp,2154)
  
  # Initialisation
  nearest_polygon <- c()
  pourcent <- ceiling(nrow(cor_carre_obs)/100)
  j <- c()

  #print("START PHASE 1/2")

  for(i in 1:nrow(cor_carre_obs)){
    if(i%%(pourcent*10) == 0 && i!=nrow(cor_carre_obs)){
      #print(paste(ceiling(i/pourcent),"%", sep =""))
      save(nearest_polygon, file = paste("Test_",ceiling(i/pourcent)/10,".Rda",sep=""))
      rm(nearest_polygon)
      nearest_polygon <- c()
      j <- c(j,ceiling(i/pourcent)/10)
    }
    carre <- cor_carre_obs[i,]$cor_carre
    indice <- st_nearest_feature(carre,polygon_Lambert)
    linestring <- st_nearest_points(carre,polygon_Lambert[indice,])
    point_pol <- st_cast(linestring, "POINT")[2]
    distance <- st_distance(carre,point_pol)
    units(distance) <- with(ud_units, km)
    distance <- round(distance,2)
    row = st_sf(point_pol=point_pol, id_polygon=polygon_Lambert[indice,]$id_polygone, polygon=polygon_Lambert[indice,]$geometry, area=polygon_Lambert[indice,]$area, length=polygon_Lambert[indice,]$lenght, distance=distance)
    nearest_polygon <- rbind(nearest_polygon,row)
    if(i == nrow(cor_carre_obs)){
      #print("100%")
      save(nearest_polygon, file = paste("Test_",10,".Rda",sep=""))
      rm(nearest_polygon)
      j <- c(j,10)
    }
  }
  
  #print("END PHASE 1/2")
  
  rlist <- c()
  #print("START PHASE 2/2")
  
  nearest_polygon_f <- c()
  
  for(i in j){
    #print(paste(i*10,"%", sep =""))
    load(paste("Test_",i,".Rda",sep=""))
    rlist <- c(rlist,(paste("Test_",i,".Rda",sep=""))) 
    nearest_polygon_f = rbind(nearest_polygon_f,nearest_polygon)
  }
  
  
  #print("END PHASE 2/2")
  file.remove(rlist)
  rm(rlist)
  
  

  # Merging tables
  data <- cbind(cor_carre_obs,nearest_polygon_f)
  
  # Delete lines with edges without coordinates
  datatable <- subset(data,id_polygon != "NA")
  return(datatable)
}
```



# Function
# Get the closest point of the nearest polygon for an IN observation point
# -> Need to cast POLYGON in LINESTRING
```{r}
nearest_points_int <- function(cor_carre_obs,area_sp){
  
  if(is.null(nrow(cor_carre_obs))){
    dataframe_vide <- c()
    return(dataframe_vide)
  }
  
  # Conversion to Lambert coordinates
  polygon_Lambert <- st_transform(area_sp,2154)
  
  # Initialisation
  nearest_polygon <- c()
  pourcent <- ceiling(nrow(cor_carre_obs)/100)
  j <- c()

  #print("START PHASE 1/2")

  for(i in 1:nrow(cor_carre_obs)){
    if(i%%(pourcent*10) == 0 && i!=nrow(cor_carre_obs)){
      #print(paste(ceiling(i/pourcent),"%", sep =""))
      save(nearest_polygon, file = paste("Test_",ceiling(i/pourcent)/10,".Rda",sep=""))
      rm(nearest_polygon)
      nearest_polygon <- c()
      j <- c(j,ceiling(i/pourcent)/10)
    }
    carre <- cor_carre_obs[i,]$cor_carre
    indice <- st_nearest_feature(carre,polygon_Lambert)
    test <- try(polygon <- st_geometry(polygon_Lambert[indice,])  %>% st_cast(to = 'LINESTRING'), silent = T)
    if(class(test)[1] == "try-error"){
      print("Error cast MULTIPOLYGON TO LINESTRING")
    }
    linestring <- st_nearest_points(carre,polygon)
    point_pol <- st_cast(linestring, "POINT")[2]
    distance <- st_distance(carre,polygon)
    units(distance) <- with(ud_units, km)
    distance <- round(distance,2)
    row = st_sf(point_pol=point_pol, id_polygon=polygon_Lambert[indice,]$id_polygone, polygon=polygon_Lambert[indice,]$geometry, area=polygon_Lambert[indice,]$area, length=polygon_Lambert[indice,]$lenght, distance=distance)
    nearest_polygon <- rbind(nearest_polygon,row)
    if(i == nrow(cor_carre_obs)){
      #print("100%")
      save(nearest_polygon, file = paste("Test_",10,".Rda",sep=""))
      rm(nearest_polygon)
      j <- c(j,10)
    }
  }
  
  #print("END PHASE 1/2")
  
  rlist <- c()
  #print("START PHASE 2/2")
  
  nearest_polygon_f <- c()
  
  for(i in j){
    #print(paste(i*10,"%", sep =""))
    load(paste("Test_",i,".Rda",sep=""))
    rlist <- c(rlist,(paste("Test_",i,".Rda",sep=""))) 
    nearest_polygon_f = rbind(nearest_polygon_f,nearest_polygon)
  }
  
  
  #print("END PHASE 2/2")
  file.remove(rlist)
  rm(rlist)

  # Merging tables
  data <- cbind(cor_carre_obs,nearest_polygon_f)
  
  # Delete lines with edges without coordinates
  datatable <- subset(data,id_polygon != "NA")
  return(datatable)
}
```


# Function
# Get the X,Y cordinates of POL & OBS point
```{r}
# Give all the coordinates in Lambert
get_cordinates <- function(data){
  
  cor_obs <- as.data.frame(st_coordinates(data$cor_carre))
  colnames(cor_obs)[1] <- "X_obs"
  colnames(cor_obs)[2] <- "Y_obs"
  
  cor_pol <- as.data.frame(st_coordinates(data$point_pol))
  colnames(cor_pol)[1] <- "X_pol"
  colnames(cor_pol)[2] <- "Y_pol"
  
  Segment <- cbind(cor_obs,cor_pol,data$distance)
  colnames(Segment)[5] <- "distance"
  
  return(Segment)
}
```



# Function 
# return the horizontal distance and vertical distance in absolute value
```{r}
distance_h_v_absolue <- function(Segment){
  
  distance_horizontale <- data.frame()
  distance_verticale <- data.frame()
  signe_horizontal <- c()
  signe_verticale <- c()
  
  for(i in 1:nrow(Segment)){
    dist_h = round((Segment[i,]$X_obs - Segment[i,]$X_pol)/1000,2)
    if(dist_h > 0){
      signe_h = 1
    }
    else{
      signe_h = -1
      dist_h = dist_h*(-1)
    }
    signe_horizontal <- c(signe_horizontal,signe_h)
    distance_horizontale <- rbind(distance_horizontale,dist_h)
    dist_v = round((Segment[i,]$Y_obs - Segment[i,]$Y_pol)/1000,2)
    if(dist_v > 0){
      signe_v = 1
    }
    else{
      signe_v = -1
      dist_v = dist_v*(-1)
    }
    signe_verticale <- c(signe_verticale,signe_v)
    
    distance_verticale <- rbind(distance_verticale,dist_v)
  }

  dataframe <- cbind(distance_horizontale,signe_horizontal,distance_verticale,signe_verticale)
  colnames(dataframe)[1] <- "distance_horizontale"
  colnames(dataframe)[2] <- "signe_horizontale"
  colnames(dataframe)[3] <- "distance_verticale"
  colnames(dataframe)[4] <- "signe_verticale"
  
  dataframe[,1] <- set_units(dataframe[,1], km)
  dataframe[,3] <- set_units(dataframe[,3], km)
  
  return(dataframe)
} 
```

  
# Run
# Take approximately 3 hours
```{r}
code_sp <- unique(all_obs_STOC$code_sp)
count = 0

for(species in code_sp){
  count <- count+1
  cat("\n",count,"\n",species,"\n")
  area_sp <- area(species,espece_BL_STOC,BL_all_polygons_F)
  if(nrow(area_sp) > 0){
    obs_sp <- observation(species,all_obs_STOC)
    if(nrow(obs_sp) > 0){
      carre_obs_sp <- get_carre(obs_sp)
      carre_int <- position_car_pol(carre_obs_sp,area_sp,1)
      carre_ext <- position_car_pol(carre_obs_sp,area_sp,0)
      nearest_pi <- nearest_points_int(carre_int,area_sp)
      nearest_pe <- nearest_points_ext(carre_ext,area_sp)
      nearest_points <- rbind(nearest_pi,nearest_pe)
      nearest_points <- nearest_points[,-9]
      obs_sp <- obs_sp[,-(6:7)]
      dataset <- inner_join(obs_sp,nearest_points)
      cordinates_h_v <- get_cordinates(dataset)
      distance_h_v <- distance_h_v_absolue(cordinates_h_v)
      species_dataset <- cbind(dataset,distance_h_v)
      save(species_dataset, file = paste("Final_",species,"_dataset.Rda",sep=""))
    }
    else{
      print("No STOC observations")
    }
  }
  else{
    print("No Birdlife polygons")
  }
  
}


print("START PHASE 2/2")
rlist <- c()
all_species_dataset <- data.frame() 
sp <- c()
count = 0

for(species in code_sp){
  species_dataset = 0
  try(load(paste("Final_",species,"_dataset.Rda",sep="")),silent = T)
  if(length(species_dataset) > 2){
    count <- count+1
    print(count)
    print(species)
    rlist <- c(rlist,(paste("Final_",species,"_dataset.Rda",sep=""))) 
    all_species_dataset = rbind(all_species_dataset,species_dataset)
  }
  else{
    print(species)
    sp <- c(sp,species)
  }
}
  
print("END PHASE 2/2")
# if you want to remove all the species file:
#file.remove(rlist)
```


# Save
```{r}
save_path = "/Users/Schneider/STAGE_STOC_MNHN/Donnees/"
save(all_species_dataset, file = paste(save_path,"All_species_dataset.Rda", sep = ""))
```


# Cleaning
```{r}
rm(list=ls())
```
