---
title: "Teletravail 3"
author: "Thomas Schneider--Bruchon"
date: "30/03/2020"
output: html_document
---

# Map Point-Polygon Distance

# Environnement 
```{r}
rm(list=ls())
library(sf)
library(dplyr)
library(units)
library(lubridate)
```


# Variables
```{r}
species = "Saxicola rubetra"
path = "/Users/Schneider/STAGE_STOC_MNHN/Donnees"
species = "Numenius arquata"
passage = 'all'
year = "all"
```

# Loading data table 
```{r}
load(paste(path,"/observation_STOC.Rda", sep = ""))

load(paste(path,"/espece_BL_STOC.Rda", sep = ""))

load(paste(path,"/Point_obs_STOC.Rda", sep = ""))

load(paste(path,"/BL_all_polygons_F.Rda", sep = ""))

load(paste(path,"/inventaire_STOC.Rda", sep = ""))
```


# Function
# Recovery of the species area #
```{r}
area <- function(species, area_table){
    
    #get the species distribution area
    try(area <- area_table %>%
    filter(SCINAME %in% species), silent = T)
    
    if(length(area)<1){
      return(paste("Error,",species,"n'est pas dans la table de comparaison"))
    }
  
    #get only resident and breeding season
    try(good_area <- area %>%
    filter(SEASONAL %in% c(1,2)), silent = T)
  
    return(good_area)
}
```



# Recovery of the points coordinates
```{r}
cordinates <- function(species, observation_data, inventaire_data, table_comparaison, table_point, passage,year){
    
    #get species code
    try(row <- table_comparaison %>%
    filter(SCINAME %in% species), silent = T)
    sp_code <- as.character(row$pk_species)
    
    if(length(sp_code)<1){
      return(paste("Error,",species,"n'est pas dans la table de comparaison"))
    }
    
    
    #get the points id
    try(row <- observation_data %>%
    filter(espece %in% sp_code), silent = T)
    id_point_ob <- row$id_point
    
    #get passage
    try(row <- inventaire_data %>%
    filter(id_point_ob %in% id_inventaire_ob), silent = T)
    passage_ob <- row$passage
    id_point_inv <- row$id_point
    date <- year(as.Date(row$date))
    
    id_point_p1 <- c()
    id_point_p2 <- c()
    
    for(i in 1:length(id_point_inv)){
      if(year == "all"){
        if(passage_ob[i] == 1){
          id_point_p1 <- c(id_point_p1, as.character(id_point_inv[i])) 
        }
        if(passage_ob[i] == 2){
          id_point_p2 <- c(id_point_p2, as.character(id_point_inv[i])) 
        }
      }
      else{
        if(date[i] == year){
          if(passage_ob[i] == 1){
            id_point_p1 <- c(id_point_p1, as.character(id_point_inv[i])) 
          }
          if(passage_ob[i] == 2){
            id_point_p2 <- c(id_point_p2, as.character(id_point_inv[i])) 
          }
        }
      }
    }
    
    cor_pt_1 <- NULL
    cor_pt_2 <- NULL
    
    #get the points cordinates
    if(passage == "all"){
      try(cor_pt_1 <- table_point %>%
      filter(pk_point %in% id_point_p1))
      
      try(cor_pt_2 <- table_point %>%
      filter(pk_point %in% id_point_p2))
    }
    
    else{
      if(passage == 1){
        try(cor_pt_1 <- table_point %>%
        filter(pk_point %in% id_point_p1))
        cor_pt_2 <- NULL
      }
      if(passage == 2){
        try(cor_pt_2 <- table_point %>%
        filter(pk_point %in% id_point_p2))
        cor_pt_1 = NULL
      }
    }
    
    dataframe <- rbind(cor_pt_1,cor_pt_2)
    return(dataframe)
}
```



# Function
# returning position nearest polygon point
```{r}

position_pt_pol <- function(Cordinates,polygon_Lambert){
  
  #Recovery points in polygons and points outside polygons
  point_dans_polygon <- c()
  point_ext_polygon <- c()
  
  for(i in 1:nrow(Cordinates)){
    point <- Cordinates[i,]$geom93
    if(st_within(point,polygon_Lambert) %>% lengths > 0){
      row = st_sf(id_point_obs=Cordinates[i,]$pk_point, point_obs=point)
      point_dans_polygon <- rbind(point_dans_polygon,row)
    }
    else{
      row = st_sf(id_point_obs=Cordinates[i,]$pk_point, point_obs=point)
      point_ext_polygon <-  rbind(point_ext_polygon,row)
    }
  }
  
  if(length(point_ext_polygon) == 0){
    return(0)
  }
  
  #Closest polygon point recovery (FOR EXT POL POINTS)
  nearest_polygon <- c()

  for(i in 1:nrow(point_ext_polygon)){
    point <- point_ext_polygon[i,]$point
    indice <- st_nearest_feature(point,polygon_Lambert)
    linestring <- st_nearest_points(point,polygon_Lambert[indice,])
    point_pol <- st_cast(linestring, "POINT")[2]
    distance <- st_distance(point,point_pol)
    units(distance) <- with(ud_units, km)
    distance <- round(distance,2)
    row = st_sf(point_pol=point_pol, polygon=polygon_Lambert[indice,]$geometry, area=polygon_Lambert[indice,]$area, length=polygon_Lambert[indice,]$lenght, distance=distance)
    nearest_polygon <- rbind(nearest_polygon,row)
  }
  
  #Merging tables
  data <- cbind(point_ext_polygon,nearest_polygon)
  
  
  #Verification no empty Geometry
  list_supp <- c()
  for(i in 1:nrow(data)){
    cor_obs <- data[i,]$point_obs
    cor_pol <- data[i,]$point_pol
    if(is.na(st_dimension(st_sfc(cor_obs))) || is.na(st_dimension(st_sfc(cor_obs)))){
      list_supp <- c(list_supp,i)
    }
  }
  list_supp <- sort(list_supp, decreasing = T)
  for(i in list_supp){
    data <- data[-i,]
    print(paste("Row",i,"was deleted because it contained GEOMETRYCOLLECTION EMPTY", sep =" "))
  }

  return(data)
}
```


# Dataframe Construction  
```{r}

# Gives coordinates in WGS to make the map
get_Segment_WGS <- function(data){
  cor_obs <- st_transform(data$point_obs,4326)
  cor_obs_WGS <- as.data.frame(st_coordinates(cor_obs))
  colnames(cor_obs_WGS)[1] <- "X_obs"
  colnames(cor_obs_WGS)[2] <- "Y_obs"
  
  cor_pol <- st_transform(data$point_pol,4326)
  cor_pol_WGS <- as.data.frame(st_coordinates(cor_pol))
  colnames(cor_pol_WGS)[1] <- "X_pol"
  colnames(cor_pol_WGS)[2] <- "Y_pol"
  
  Segment <- cbind(cor_obs_WGS,cor_pol_WGS,data$distance)
  colnames(Segment)[5] <- "distance"
  
  return(Segment)
}


# Gives the coordinates in Lambert
get_Segment_Lambert <- function(data){
  
  cor_obs <- as.data.frame(st_coordinates(data$point_obs))
  colnames(cor_obs)[1] <- "X_obs"
  colnames(cor_obs)[2] <- "Y_obs"
  
  cor_pol <- as.data.frame(st_coordinates(data$point_pol))
  colnames(cor_pol)[1] <- "X_pol"
  colnames(cor_pol)[2] <- "Y_pol"
  
  Segment <- cbind(cor_obs,cor_pol,data$distance)
  colnames(Segment)[5] <- "distance"
  
  return(Segment)
}


```


# Execution function 1
# Return a datatable which contains the cordinates of the points outside of the polygons for one species (WGS cordinates)
```{r}
proximite_WGS <- function(species,BL_all_polygons_F,observation_STOC,inventaire_stoc,espece_BL_STOC,Point_obs_STOC,passage,year){
  
  Area <- area(species,BL_all_polygons_F)
  Cordinates <- cordinates(species,observation_STOC,inventaire_stoc,espece_BL_STOC,Point_obs_STOC,passage,year)
  
  if(nrow(Cordinates) == 0){
    print(paste("No observation for the species",species,"in year",year, sep=" "))
    return(0)
  }
  
  #Area$geometry WGS84 -> lambert
  polygon_Lambert <- st_transform(Area,2154)
  
  data <- position_pt_pol(Cordinates,polygon_Lambert) 

  if(is.null(nrow(data))){
    print(paste("There was no observation point outside polygon for the species",species,"in year",year, sep=" "))
    return(0)
  }
  
  Segment <- get_Segment_WGS(data)
  
  return(Segment)
}
```


# Execution function 2
# Return a datatable which contains the cordinates of the points outside of the polygons for one species (Lambert cordinates)
```{r}
proximite_Lambert <- function(species,BL_all_polygons_F,observation_STOC,inventaire_stoc,espece_BL_STOC,Point_obs_STOC,passage,year){
  
  Area <- area(species,BL_all_polygons_F)
  Cordinates <- cordinates(species,observation_STOC,inventaire_stoc,espece_BL_STOC,Point_obs_STOC,passage,year)
  
  if(nrow(Cordinates) == 0){
    print(paste("No observation for the species",species,"in year",year, sep=" "))
    return(0)
  }
  
  # Area$geometry WGS84 -> lambert
  polygon_Lambert <- st_transform(Area,2154)
  
  data <- position_pt_pol(Cordinates,polygon_Lambert) 

  if(is.null(nrow(data))){
    print(paste("There was no observation point outside polygon for the species",species,"in year",year, sep=" "))
    return(0)
  }
  
  Segment <- get_Segment_Lambert(data)
  
  return(Segment)
}
```


# Execution function 3
# Return a data table which contains the points outside of the polygons and informations about these polygons
```{r}
all_data <- function(species,BL_all_polygons_F,observation_STOC,inventaire_stoc,espece_BL_STOC,Point_obs_STOC,passage,year){
  
  Area <- area(species,BL_all_polygons_F)
  Cordinates <- cordinates(species,observation_STOC,inventaire_stoc,espece_BL_STOC,Point_obs_STOC,passage,year)
  
  if(nrow(Cordinates) == 0){
    print(paste("No observation for the species",species,"in year",year, sep=" "))
    return(0)
  }
  
  #Area$geometry WGS84 -> lambert
  polygon_Lambert <- st_transform(Area,2154)
  
  data <- position_pt_pol(Cordinates,polygon_Lambert)
  
  if(is.null(nrow(data))){
    print(paste("There was no observation point outside polygon for the species",species,"in year",year, sep=" "))
    return(0)
  }
  
  return(data)
}
```



# Execution
```{r}
Segment_WGS <- proximite_WGS(species,BL_all_polygons_F,observation_STOC,inventaire_stoc,espece_BL_STOC,Point_obs_STOC,passage,year)

Segment_Lambert <- proximite_Lambert(species,BL_all_polygons_F,observation_STOC,inventaire_stoc,espece_BL_STOC,Point_obs_STOC,passage,year)

data <- all_data(species,BL_all_polygons_F,observation_STOC,inventaire_stoc,espece_BL_STOC,Point_obs_STOC,passage,year)
```



# Map creation 
```{r}
species = "Sylvia melanocephala"
monde <- map_data("world")
Area <- area(species,BL_all_polygons_F)
Segment <- proximite_WGS(species,BL_all_polygons_F,observation_STOC,inventaire_stoc,espece_BL_STOC,Point_obs_STOC,passage,year)
Cordinates <- cordinates(species,observation_STOC,inventaire_stoc,espece_BL_STOC,Point_obs_STOC,passage,year)
description <- c("1" = "Resident", "2"= "Breeding Season", "3" = "Non-breeding Season", "4" = "Passage")
cols <- c("1" = "darkmagenta", "2"= "firebrick1","3" = "deepskyblue2", "4" = "gold")

map <- ggplot()+
      geom_polygon(data=monde,aes(x=long,y=lat,group=group),fill="papayawhip",col="black")+
      geom_sf(data = Area, aes(fill = factor(SEASONAL)), alpha=0.4)+
      geom_point(data=Cordinates,aes(x=longitude_wgs84, y=latitude_wgs84),col = "gray0", size = 1)+
      geom_segment(data = Segment, aes(x = X_obs, xend = X_pol, y = Y_obs, yend = Y_pol), col = "green4", size = 1)+
      geom_text(data = Segment %>% filter(as.numeric(distance)>31), aes(x = X_obs, y = Y_obs, label=distance), nudge_y = -0.2, check_overlap = TRUE, size = 2, col = "red")+
      scale_fill_manual(values = cols, name = NULL, labels = description)+
      coord_sf(xlim = c(-10,10), ylim = c(40,55))+
      guides(alpha = "none")+
      theme_void()+
      labs(title = paste("Aire de répartition de",species,year), caption = "based on BirdLife & STOC-eps data")+
      theme(panel.background = element_rect(fill ="slategray1"), plot.title =element_text(color = "#063376", hjust = 0.5, vjust = 3), plot.caption = element_text(face = "bold", vjust = 1), legend.margin = margin(0,0,0,1, "cm"))

map
```

# Save
```{r}
#Save the last created map
save_path = "/Users/Schneider/STAGE_STOC_MNHN/Cartes/"
ggsave(paste(save_path,species,"_",passage,"_",year,".png", sep = ""))
```
